$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-05 paper/paper.tex
on congruence

\begin{nitheorem}[Subject congruence]
  \label{thm:subject-congruence}
  Applying structural congruence rules to a well-typed process preserves its well-typedness.
  That is, if $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

\todo{considering that Equals is in the appendix, you cannot do induction on something that is not defined in the main text. Bring the proof in the appendix. Maybe briefly comment on the proof here}
\begin{proof}
  By induction on \textsc{Equals}.
  The base cases under $\constr{struct}$ (and their symmetric variants under $\constr{sym}$) are the nontrivial ones.
  For those, we proceed by induction on \textsc{StructCong}:
  \begin{itemize}
    \item
      The case $\constr{comp-assoc}$ is trivial: leftover typing is naturally associative.
    \item
      In the $\constr{comp-sym}$ case for a process $\comp{P}{Q}$, we need framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of of $Q$ to the one of $P$.
      The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
    \item
      The case $\constr{comp-end}$ is trivial: the typing rule for a process $\PO$ has the same input and output contexts.
    \item
      In the $\constr{scope-end}$ case, we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
    \item
      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
      We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with typing derivation on $Q$ wrapped in $\constr{chan}$.
      In the reverse direction, we weaken (\autoref{thm:weakening}) $P$ so that it accepts the additional variable.
      We then use \autoref{lm:lower-lift} to show that lowering and then lifting $P$ results in $P$.
    \item
      In the $\constr{scope-comm}$ case, we use swapping (\autoref{thm:swapping}).
      In the reverse direction, we use swapping and then \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.
  \end{itemize}
\end{proof}

>>>>

Subject congruence theorem states that applying structural congruence rules to a well-typed process preserves its well-typedness.
\begin{nitheorem}[Subject congruence]
  \label{thm:subject-congruence}
  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

\begin{proof}[Proof (Sketch)]
  The proof is induction on \textsc{Equals} $\eq{}$.
  Here we will only consider the base cases for $\constr{struct}$ (and their symmetric variants under $\constr{sym}$), which are nontrivial. Full proof in \cite{Zalakain2020Agda}.
  We proceed by induction on \textsc{StructCong} $\equiv$:
  \begin{itemize}
    \item
      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
    \item
      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
      The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
    \item
      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
    \item
      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
    \item
      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
      We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with typing derivation on $Q$ wrapped in $\constr{chan}$.
      In the reverse direction, we weaken (\autoref{thm:weakening}) $P$ so that it accepts the additional variable.
      We then use \autoref{lm:lower-lift} to show that lowering and then lifting $P$ results in $P$.
    \item
      Case $\constr{scope-comm}$: in one direction we use swapping (\autoref{thm:swapping}).
      In the reverse direction, we use swapping and then \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.
  \end{itemize}
\end{proof}
